Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-10-22 text/supp.tex
Remove extra space in SOM
Edits to SOM

\item We filled in single recorded values of zero with the lowest non-zero value
  in the time series \citep[following][]{brook2006a}. This assumes that single
  values of zero result from abundance being low enough that censusing missed
  present individuals. We turned multiple zero values in a row into \texttt{NA}
  values. This implies that multiple zero values were either censusing errors or
  caused by emigration. Regardless, our population models were fitted on
  a multiplicative (log) scale and so could not account for zero abundance.

@@

\item Finally, we removed GPDD Main ID \texttt{20531}, which we noticed was
  a duplicate of \texttt{10139} (a heron population).

@@

For the Gompertz model, our weakly-informative priors (Fig.~\ref{fig:priors}) were:
\begin{align*}
b &\sim \mathrm{Uniform}(-1, 2)\\
\lambda &\sim \mathrm{Normal}(0, 10^2)\\
\sigma &\sim \mathrm{Half\mhyphen Cauchy} (0, 2.5)\\
\nu &\sim \mathrm{Truncated\mhyphen Exponential}(0.01, \mathrm{min.} = 2). \end{align*}
Our prior on $b$ was uninformative between values of $-1$ and $2$. We would not
expect values of $b$ with levels of inverse density dependence as low as $-1$,
nor would we expect values above $1$. We allowed values of $b$ above $1$ to
allow for somewhat non-stationary time series of growth rates. The estimates of
$b$ were well within these bounds. Our prior on $\lambda$ was very weakly
informative within the range of expected values for population growth and is
similar to priors suggested by \citet{gelman2008d} for intercepts of regression
models. Our prior on $\sigma$ follows \citet{gelman2006c} and
\citet{gelman2008d} and is based on our expected range of the value in nature
from previous studies \citep[e.g.][]{connors2014}. In our testing of a subsample
of populations, our parameter estimates were not qualitatively changed by
switching to an uninformative uniform prior on $\sigma$, but the weakly
informative prior substantially sped up chain convergence.

Our prior on $\nu$ was based on \citet{fernandez1998}; they chose a more
informative exponential rate parameter of $0.1$. We chose a less informative
rate parameter of $0.01$ and truncated the distribution at $2$, since at $\nu
< 2$ the variance of the t distribution is undefined. This prior gives only
a $7.7$\% probability $\nu < 10$ but constrains the sampling sufficiently to
avoid wandering off towards infinity --- above approximately $\nu = 20$ the
t distribution is so similar to the normal distribution
(Fig.~\ref{fig:didactic}) that time series of the length considered here are
unlikely to be informative about the precise value of $\nu$. In the
scenario where the data are not heavy tailed (e.g.~Fig.~\ref{fig:didactic}e, h)
the posterior will approximately match the prior (median $= 71$, mean $= 102$)
and not be flagged as likely heavy tailed using the metrics we used in our paper
(e.g.~Pr$(\nu < 10) > 0.5$).

We fitted our models with Stan 2.4.0 \citep{stan-manual2014}, and R 3.1.1
\citep{r2014}. We began with four chains and $2000$ iterations, discarding the
first $1000$ as warm up (i.e.~4000 total samples). If $\hat{R}$ (a measure of
chain convergence) was greater than $1.05$ for any parameter or the minimum
effective sample size, $n_\mathrm{eff}$, (a measure of the effective number of
uncorrelated samples) for any parameter was less than $200$, we doubled both the
total iterations and warm up period and sampled from the model again. These
thresholds are in excess of the minimums recommended by \citet{gelman2006a} of
$\hat{R} < 1.1$ and effective sample size $> 100$ for reliable point estimates
and confidence intervals. In the majority of cases our minimum thresholds were
greatly exceeded. We continued this procedure up to $8000$ iterations ($16000$
total samples) by which all chains were deemed to have sufficiently converged.

@@

We performed two types of simulation testing. First, we tested how easily the
Student-t distribution $\nu$ parameter could be recovered given different true
values of $\nu$ and sample sizes. Second, we tested the ability of the
heavy-tailed Gompertz model to obtain unbiased parameter estimates of $\nu$
given that a set of process deviations was provided in which the ``effective
$\nu$'' value was close to the true $\nu$ value.

@@

First, we tested the ability to estimate $\nu$ given different true values of
$\nu$ and sample sizes. We took stochastic draws from t distributions with
different $\nu$ values ($\nu = 3, 5, 10,$ and $10^6$ [$\approx$ normal]), with
central tendency parameters of $0$, and scale parameters of $1$. We started with
$1600$ stochastic draws and then fitted the models again at the first $800, 400,
200, 100, 50,$ and $25$ draws. Each time we recorded the posterior samples of
$\nu$.

We found that we could consistently and precisely recover median posterior
estimates of $\nu$ near the true value of $\nu$ with large samples ($\ge 200$)
(Fig.~\ref{fig:sim-gompertz} upper panels). At smaller samples we could still
usually distinguish ``heavy'' from ``not-heavy'' tails, but the model tended to
underestimate how heavy the tails were. At the same time, the model tended to
overestimate how large the scale parameter was (Fig.~\ref{fig:sim-gompertz}
lower panels).

@@

In the second part of our simulation testing, we tested the ability of the
heavy-tailed Gompertz model to obtain unbiased parameter estimates given that
process deviations were provided in which our simpler model (previous
section) could accurately estimate the value of $\nu$. To generate these process
deviations, we repeatedly generated candidate process deviations and estimated the
central tendency, scale, and $\nu$ values each time. We recorded when
$\hat{\nu}$ was within $0.2$ CVs (coefficient of variations) of the true $\nu$
value and used this set of random seed values in our Gompertz simulation. We
then fitted our Gompertz models to the simulated datasets with all parameters
(except $\nu$) set near the median values estimated in the GPDD. We repeated
this with $50$ and $100$ samples without observation error, $50$ samples with
observation error ($\sigma_\mathrm{obs} = 0.2$), and $50$ samples with the same
observation error and a Gompertz model that allowed for correctly specified
observation error magnitude.

@@

When converting the posterior distributions of $\nu$ into Pr($\nu < 10$) the
models performed well. Without observation error, and using a probability of
$0.5$ as a threshold, the model would have miscategorized only one of $40$
simulations at $\nu = 5$ across a sample size of $50$ or $100$ data points
(Fig.~\ref{fig:sim-prob}, upper panels). The model would have correctly
categorized all cases where the process noise was not heavy tailed (indicated as
``$\nu =$ infinity'' in Fig.~\ref{fig:sim-prob}) and all cases where $\nu = 3$.
With $0.2$ standard deviations of observation error, the model still categorized
XX\% of cases as heavy tailed when $\nu = 5$ and all cases where $\nu = 3$.
Allowing for observation error made little improvement to the detection of heavy
tails (Fig.~\ref{fig:sim-prob}, lower-left vs.\ lower-right panel). Therefore,
we chose to focus on the simpler model without observation error in the main
text.

@@

We fit four alternative population models to the time-series data to check how
they would influence our conclusions. Our alternative models allowed for
autocorrelation in the residuals, assumed no density dependence, allowed for
observation error, or assumed a Ricker-logistic functional form. The ranges of
percentages of black swans by taxonomic class cited in the abstract are based on
lower and upper limits across our main Gompertz model and these four alternative
models.

@@

We fit a simplified version of the Gompertz model in which the density
dependence parameter $b$ was fixed at $1$ (density independent). This is
equivalent to fitting a random walk model (with drift) to the $\ln$ abundances
or assuming the growth rates are stationary. The model was as follows:

\begin{align*}
x_t &= \lambda + x_{t-1} + \epsilon_t\\
\epsilon &\sim \mathrm{Student\mhyphen t}(\nu, \epsilon, \sigma).
\end{align*}
We fit this model for three reasons: (1) it is computationally simpler and so
provides a check that our more complicated full Gompertz model was obtaining
ballpark reasonable estimates of $\nu$, (2) it provides a test of whether
density dependence was systematically affecting our perception of heavy tails,
(3) it matches how some previous authors have modelled heavy tails without
accounting for density dependence \citep{segura2013}.

@@

Observation error can bias parameter estimates \citep[e.g.][]{knape2012} and is
known to affect the ability to detect extreme events \citep{ward2007}. However,
simultaneously estimating observation and process noise is a challenging
problem and is known to result in identifiability issues with the Gompertz
population model \citep{knape2008}. Furthermore, our model estimates an
additional parameter --- the shape of the process deviation tails ---
potentially making identifiability and computational issues even greater.
Therefore, we considered a version of the base Gompertz model that allowed for
a fixed level of observation error:
\begin{align*}
U_t &= \lambda + b U_{t-1} + \epsilon_t\\
x_t &\sim \mathrm{Normal}(U_t, \sigma_\mathrm{obs})\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma_\mathrm{proc}),
\end{align*}
where $U$ represents the unobserved state vector, and $\sigma_\mathrm{obs}$
represents the standard deviation of observation error (on a log scale). We set
$\sigma_\mathrm{obs}$ to $0.2$, which represents the upper end of values often
used in simulation analyses \citep[e.g.][]{valpine2002, thorson2014b}.

@@

We also fitted a Ricker-logistic model:
\begin{align*}
x_t &= x_{t-1} + r_{\mathrm{max}}\left(1 - \frac{N_{t-1}}{K}\right) + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma),
\end{align*}
where $r_\mathrm{max}$ represents the maximum population growth rate that is
obtained when $N$ (abundance) $= 0$. The parameter $K$ represents the carrying
capacity and, as before, $x_t$ represents the $\ln$ transformed abundance at
time $t$. The Ricker-logistic model assumes a linear decrease in population
growth rate ($x_t - x_{t-1}$) with increases in abundance ($N_t$). In contrast,
the Gompertz model assumes a linear decrease in population growth rate with
increases in $\ln$ abundance ($x_t$) (REF).

@@

We fitted a multilevel beta regression model to the predicted probability of
heavy tails, Pr($\nu < 10$), to investigate potential covariates of heavy-tailed
dynamics. The beta distribution is useful when the response data range on
a continuous scale between zero and one. We used a logit link function as is
typically used in logistic regression. The model was as follows:
\begin{align*}
\mathrm{Pr}(\nu_i < 0.5) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + X_i \beta),
  \: \text{for } i = 1, \dots, 617\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,\\
\end{align*}
where $A$ and $B$ represent the beta distribution shape parameters; $\mu_i$
represents the predicted value for population $i$, class $j$, order $k$, and
species $l$; $\phi_\mathrm{disp}$ represents the dispersion parameter; and $X_i$
represents a vector of predictors for population $i$ with associated $\beta$
parameters. The intercepts are allowed to vary from the overall intercept
$\alpha$ by taxonomic classes ($\alpha^\mathrm{class}_j$), taxonomic orders
($\alpha^\mathrm{order}_k$), and species ($\alpha^\mathrm{species}_l$) with
standard deviations $\sigma_{\alpha \; \mathrm{class}}$, $\sigma_{\alpha \;
  \mathrm{order}}$, and $\sigma_{\alpha \; \mathrm{species}}$. Where possible,
we also allowed for error distributions around the predictors by incorporating
the standard deviation of the posterior samples for the Gompertz parameters
$\lambda$, $b$, and $\sigma$ around the mean point value (not shown in the above
equation). We log transformed $\sigma$, time-series length, and lifespan to
match the way they are visually represented in Fig.~\ref{fig:correlates}. All
input variables were standardized by subtracting their mean and dividing by two
standard deviations \citep{gelman2008c} to make their coefficients comparable in
magnitude.

@@

To derive taxonomic-order-level estimates of the probability of heavy tails
accounting for time-series length (Fig \ref{fig:order-estimates}), we fitted
a separate multilevel model with the same structure but with only time-series
length as a predictor. (In this case, we did not want to control for intrinsic
population characteristics such as density dependence.) We obtained order-level
estimates by adding the posteriors for $\alpha$, $\alpha^\mathrm{class}_j$, and
$\alpha^\mathrm{order}_k$.

>>>>

\item We filled in single recorded values of zero with the lowest non-zero value
  in the time series \citep[following][]{brook2006a}. This assumes that single
  values of zero result from abundance being low enough that censusing missed
  present individuals. We turned multiple zero values in a row into \texttt{NA}
  values. This implies that multiple zero values were either censusing errors or
  caused by emigration. Regardless, our population models were fitted on
  a multiplicative (log) scale and so could not account for zero abundance. To
  avoid distorting the original data too strongly, we removed populations in
  which we filled in more than four zeros.

@@

\item Finally, we removed GPDD Main ID \texttt{20531}, which we noticed was
  a duplicate of \texttt{10139} (a heron population). We removed a limited
  number of populations from class Angiospermopsida and Bacillariophyceae to
  focus the taxonomy in our analysis. We also removed any populations with an
  \texttt{Unknown} taxonomic class.

@@

For the Gompertz model, our weakly-informative priors (Fig.~\ref{fig:priors}) were:
\begin{align*}
b &\sim \mathrm{Uniform}(-1, 2)\\
\lambda &\sim \mathrm{Normal}(0, 10^2)\\
\sigma &\sim \mathrm{Half\mhyphen Cauchy} (0, 2.5)\\
\nu &\sim \mathrm{Truncated\mhyphen Exponential}(0.01, \mathrm{min.} = 2). \end{align*}
Our prior on $b$ was uninformative between values of $-1$ and $2$. We would not
expect values of $b$ with levels of density dependence as low as $-1$ (very
strong inverse density dependence), nor would we generally expect values above $1$. We
allowed values of $b$ above $1$ to allow for somewhat non-stationary time
series of growth rates. The estimates of $b$ were well within these bounds. Our
prior on $\lambda$ was very weakly informative within the range of expected
values for population growth and is similar to the default priors suggested by
\citet{gelman2008d} for intercepts of regression models. Our prior on $\sigma$
follows \citet{gelman2006c} and \citet{gelman2008d} and is based on our
expected range of the value in nature from previous studies
\citep[e.g.][]{connors2014}. In our testing of a subsample of populations, our
parameter estimates were not qualitatively changed by switching to an
uninformative uniform prior on $\sigma$, but the weakly informative Half-Cauchy prior
substantially sped up chain convergence.

Our prior on $\nu$ was based on \citet{fernandez1998}. They chose a more
informative exponential rate parameter of $0.1$. We chose a less informative
rate parameter of $0.01$ and truncated the distribution at $2$, since at $\nu
< 2$ the variance of the t distribution is undefined. This prior gives only
a $7.7$\% probability $\nu < 10$ but constrains the sampling sufficiently to
avoid wandering off towards infinity --- above approximately $\nu = 20$ the
t distribution is so similar to the normal distribution
(Fig.~\ref{fig:didactic}) that time series of the length considered here are
unlikely to be informative about the precise value of $\nu$. In the
scenario where the data are not heavy tailed (e.g.~Fig.~\ref{fig:didactic}e, h)
the posterior will approximately match the prior (prior median $= 71$, mean $= 102$)
and not be flagged as likely heavy tailed using the metrics we used in our paper
(e.g.~Pr$(\nu < 10) > 0.5$).

We fitted our models with Stan 2.4.0 \citep{stan-manual2014}, and R 3.1.1
\citep{r2014}. We began with four chains and $2000$ iterations, discarding the
first $1000$ as warm up (i.e.~4000 total samples). If $\hat{R}$ (the scale
reduction factor --- a measure of chain convergence) was greater than $1.05$
for any parameter or the minimum effective sample size, $n_\mathrm{eff}$, (a
measure of the effective number of uncorrelated samples) for any parameter was
less than $200$, we doubled both the total iterations and warm up period and
sampled from the model again. These thresholds are in excess of the minimums
recommended by \citet{gelman2006a} of $\hat{R} < 1.1$ and effective sample size
$> 100$ for reliable point estimates and confidence intervals. In the majority
of cases our minimum thresholds were greatly exceeded. We continued this
procedure up to $8000$ iterations ($16000$ total samples) by which all chains
were deemed to have sufficiently converged. These chain lengths may seem low to
those familiar with software such as WinBUGS or JAGS, but the No-U-Turn
Hamiltonian Markov chain Monte Carlo Sampler in Stan generally requires far
fewer iterations to obtain equivalent effective sample sizes \citep{stan-manual2014}.

@@

We performed two types of simulation testing. First, we tested how easily the
Student-t distribution $\nu$ parameter could be recovered given different true
values of $\nu$ and different sample sizes. Second, we tested the ability of the
heavy-tailed Gompertz model to obtain unbiased parameter estimates of $\nu$
given that a set of process deviations was provided in which the effective
$\nu$ value was close to the true $\nu$ value.

@@

First, we tested the ability to estimate $\nu$ given different true values of
$\nu$ and different sample sizes. We took stochastic draws from t distributions with
different $\nu$ values ($\nu = 3, 5, 10,$ and $10^6$ [$\approx$ normal]), with
central tendency parameters of $0$, and scale parameters of $1$. We started with
$1600$ stochastic draws and then fitted the models again at the first $800, 400,
200, 100, 50,$ and $25$ draws. Each time we recorded the posterior samples of
$\nu$.

We found that we could consistently and precisely recover median posterior
estimates of $\nu$ near the true value of $\nu$ with large samples ($\ge 200$)
(Fig.~\ref{fig:sim-gompertz} upper panels). At smaller samples we could still
usually qualitatively distinguish heavy from not-heavy tails, but the
model tended to underestimate how heavy the tails were. At the same time, the
model tended to overestimate how large the scale parameter was
(Fig.~\ref{fig:sim-gompertz} lower panels).

@@

In the second part of our simulation testing, we tested the ability of the
heavy-tailed Gompertz model to obtain unbiased parameter estimates given that
process deviations were provided in which our simpler model (previous
section) could accurately estimate the value of $\nu$. To generate these process
deviations, we repeatedly drew proposed candidate process deviations and estimated the
central tendency, scale, and $\nu$ values each time. We recorded when
$\hat{\nu}$ (median of the posterior) was within $0.2$ CVs (coefficient of
variations) of the true $\nu$
value and used this set of random seed values in our Gompertz simulation. We
then fitted our Gompertz models to the simulated datasets with all parameters
(except $\nu$) set near the median values estimated in the GPDD. We repeated
this with $50$ and $100$ samples without observation error, $50$ samples with
observation error ($\sigma_\mathrm{obs} = 0.2$), and $50$ samples with the same
observation error and a Gompertz model that allowed for correctly specified
observation error magnitude.

@@

When converting the posterior distributions of $\nu$ into Pr($\nu < 10$), the
models distinguished heavy and not-heavy tails reasonably well. Without
observation error, and using a probability of $0.5$ as a threshold, the model
would have miscategorized only one of $40$ simulations at $\nu = 5$ across
a sample size of $50$ or $100$ data points (Fig.~\ref{fig:sim-prob}, upper
panels). The model would have correctly categorized all cases where the process
noise was not heavy tailed (indicated as ``$\nu =$ infinity'' in
Fig.~\ref{fig:sim-prob}) and all cases where $\nu = 3$. With $0.2$ standard
deviations of observation error, the model still categorized XX\% of cases as
heavy tailed when $\nu = 5$ and all cases where $\nu = 3$. Allowing for
observation error made little improvement to the detection of heavy tails
(Fig.~\ref{fig:sim-prob}, lower-left vs.\ lower-right panel). Therefore, we
chose to focus on the simpler model without observation error in the main
text, particularly given that the true magnitude of observation error was
unknown in the empirical data.

@@

We fit four alternative population models to the time-series data to check how
they would influence our conclusions. Our alternative models allowed for
autocorrelation in the residuals, assumed no density dependence, allowed for
observation error, or assumed a Ricker-logistic functional form. The range of
percentages of black swans by taxonomic class cited in the abstract and results
are based on lower and upper limits across our main Gompertz model and these
four alternative models.

@@

We fit a simplified version of the Gompertz model in which the density
dependence parameter $b$ was fixed at $1$ (density independent). This is
equivalent to fitting a random walk model (with drift) to the $\log$ abundances
or assuming the growth rates are drawn from a stationary distribution. The
model was as follows:
\begin{align*}
x_t &= \lambda + x_{t-1} + \epsilon_t\\
\epsilon &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma).
\end{align*}
We fit this model for three reasons: (1) it is computationally simpler and so
provides a check that our more complicated full Gompertz model was obtaining
ballpark reasonable estimates of $\nu$, (2) it provides a test of whether
density dependence was systematically affecting our perception of heavy tails,
(3) it matches how some previous authors have modelled heavy tails without
accounting for density dependence \citep{segura2013}.

@@

Observation error can bias parameter estimates \citep[e.g.][]{knape2012} and is
known to affect the ability to detect extreme events \citep{ward2007}. However,
simultaneously estimating observation and process noise is a challenging
problem and is known to result in identifiability issues with the Gompertz
population model \citep{knape2008}. Furthermore, our model estimates an
additional parameter --- the shape of the process deviation tails ---
potentially making identifiability and computational issues even greater.
Therefore, we considered a version of the base Gompertz model that allowed for
a fixed level of observation error:
\begin{align*}
U_t &= \lambda + b U_{t-1} + \epsilon_t\\
x_t &\sim \mathrm{Normal}(U_t, \sigma_\mathrm{obs}^2)\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma_\mathrm{proc}),
\end{align*}
where $U$ represents the unobserved state vector, $\sigma_\mathrm{obs}$
represents the standard deviation of observation error (on a log scale), and
$\sigma_\mathrm{proc}$ represent the process noise scale parameter. We set
$\sigma_\mathrm{obs}$ to $0.2$, which represents the upper limit of values often
used in simulation analyses \citep[e.g.][]{valpine2002, thorson2014b}.

@@

We also fitted a Ricker-logistic model:
\begin{align*}
x_t &= x_{t-1} + r_{\mathrm{max}}\left(1 - \frac{N_{t-1}}{K}\right) + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma),
\end{align*}
where $r_\mathrm{max}$ represents the theoretical maximum population growth
rate that is obtained when $N$ (abundance) $= 0$. The parameter $K$ represents
the carrying capacity and, as before, $x_t$ represents the $\log$ transformed
abundance at time $t$. The Ricker-logistic model assumes a linear decrease in
population growth rate ($x_t - x_{t-1}$) with increases in abundance ($N_t$).
In contrast, the Gompertz model assumes a linear decrease in population growth
rate with increases in \textit{log} abundance ($x_t$) (REF).

@@

We fitted a multilevel beta regression model to the predicted probability of
heavy tails, Pr($\nu < 10$), to investigate potential covariates of heavy-tailed
dynamics. The beta distribution is useful when response data range on
a continuous scale between zero and one. We used a logit link function as is
typically used in logistic regression. The model was as follows:
\begin{align*}
\mathrm{Pr}(\nu_i < 0.5) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + X_i \beta),
  \: \text{for } i = 1, \dots, 617\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,\\
\end{align*}
where $A$ and $B$ represent the beta distribution shape parameters; $\mu_i$
represents the predicted value for population $i$, class $j$, order $k$, and
species $l$; $\phi_\mathrm{disp}$ represents the dispersion parameter; and $X_i$
represents a vector of predictors (such as lifespan) for population $i$ with
associated $\beta$ coefficients. The intercepts are allowed to vary from the
overall intercept $\alpha$ by taxonomic class ($\alpha^\mathrm{class}_j$),
taxonomic order ($\alpha^\mathrm{order}_k$), and species
($\alpha^\mathrm{species}_l$) with standard deviations $\sigma_{\alpha \;
  \mathrm{class}}$, $\sigma_{\alpha \; \mathrm{order}}$, and $\sigma_{\alpha \;
  \mathrm{species}}$. Where possible, we also allowed for error distributions
around the predictors by incorporating the standard deviation of the posterior
samples for the Gompertz parameters $\lambda$, $b$, and $\sigma$ around the
mean point value as normal distributions (not shown in the above equation). We
log transformed $\sigma$, time-series length, and lifespan to match the way
they are visually represented in Fig.~\ref{fig:correlates} and to make the
relationship approximately linear on the logit-transformed response scale. All
input variables were standardized by subtracting their mean and dividing by two
standard deviations \citep{gelman2008c} to make their coefficients comparable
in magnitude.

@@

To derive taxonomic-order-level estimates of the probability of heavy tails
accounting for time-series length (Fig \ref{fig:order-estimates}), we fitted
a separate multilevel model with the same structure but with only time-series
length as a predictor. (In this case, we did not want to control for intrinsic
population characteristics such as density dependence.) Since our predictors were
centered by subtracting their mean value, we obtained order-level
estimates of the probability of heavy tails at mean time-series length by
adding the posteriors for $\alpha$, $\alpha^\mathrm{class}_j$, and
$\alpha^\mathrm{order}_k$.
